Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-07-27 paper/manuscript.tex
stylistic fixes

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig5_thermo_cyc_abs/Figure.pdf}
    \caption{\textbf{Thermodynamic cycle required for an absolute free energy calculation -- absolute free energy of binding example} The fully interacting ligand in water (\textbf{A}), has its charges turned off to pass to (\textbf{B}) followed by turning of van der Waals terms, resulting in a non-interacting ligand in water in (\textbf{C}). Restraints are used on the fully interacting ligand in the binding site of a protein or host molecule (\textbf{D}). The next step is to turn off the charges again (\textbf{E}) followed by the van der Waals interactions resulting in a non-interacting complex state (\textbf{F}). Free energyes can be computed as $\Delta G_{bind} = (\Delta G^{\mathrm{elec}}_{\mathrm{solv}}+ \Delta G^{\mathrm{VdW}}_{\mathrm{solv}})-(\Delta G^{\mathrm{elec}}_{\mathrm{bound}}+ \Delta G^{\mathrm{VdW}}_{\mathrm{bound}}$).
    }
    \label{fig:fig_absolute_thermodynamic_cycle}
\end{figure}

@@

The key consideration for choosing alchemical pathways is that the intermediate states that a given pathway produces should sample configurational ensembles that change as slowly as possible as $\vec{\lambda}$ changes, while still managing to go from the initial state to the final state as $\vec{\lambda}$ goes from 0 to 1.

@@

\begin{equation}
    U(\vec{r_{ij}},\vec{\lambda}) = 4\epsilon_{ij} \vec{\lambda} \left(\frac{1}{(\alpha(1-\vec{\lambda}) + (r_{ij}/\sigma_{ij})^6)^2} -  \frac{1}{\alpha(1-\vec{\lambda}) + (r_{ij}/\sigma_{ij})^6}\right)
    \label{eq:softcore}
\end{equation}

Where $r_{ij}$ is the distance between two particles $i$ and $j$, $\epsilon_{ij}$ and $\sigma_{ij}$ are the Lennard-Jones parameters corresponding to the interaction between particles $i$ and $j$, and $\alpha$ is a constant (0.5 is optimal for the functional form shown above).  This functional form has exactly the property we are looking for: it recovers the Lennard-Jones potential when $\vec{\lambda}=1$, and the other endpoint ($\vec{\lambda}=0$), it is exactly zero for all $r_{ij}$ everywhere, and as $\vec{\lambda}$ goes to zero, the  $\alpha(1-\vec{\lambda})$ term lowers the infinite energy in the core.  There are several different variants of the same functional form~\cite{zacharias1994separationshifted, beutler1994avoiding,pham2011identifying}, but the one  given in eq.~\ref{eq:softcore} is easy to understand and implement and fairly numerically stable. This functional form is shown in \textbf{C} and \textbf{D} of Fig.~\ref{fig:fig_what_is_lambda}.

@@

Relative calculations introduce additional choices, such as the order in which to modify nonbonded interactions.
A common process in single topology relative calculations is to first remove electrostatic interactions of any atoms which will be deleted, then modify other non-bonded interactions, then restore electrostatic interactions of any atoms which are being inserted. Although this is a simpler path to understand cognitively and can take advantage of the soft-core potential from eq~\ref{eq:softcore}, this can lead to more intermediate steps and thus be more computationally expensive.
Other schemes, such as simultaneously changing electrostatic and Lennard-Jones interactions with electrostatic ``soft-core'' potentials~\cite{steinbrecher2007nonlinear} may be implemented with fewer intermediate but could require fine-tuning of electrostatic and Lennard-Jones softcore parameters to avoid numerical instabilities. 
At the time of writting, there has not been conclusive evidence to suggest one approach is better in general than the other, so discretion should be left up to the user as to what is viable from both hardware resources, and what the simulation software supports.

A key additional consideration in choosing the alchemical pathway is the choice of spacing of intermediate states.
The spacing depends to some extent on the choice of analysis method, though states should essentially be spaced equidistant in the relevant thermodynamic length~\cite{crooks2007measuringa, sivak2012thermodynamic}.
For BAR/MBAR techniques this means that states should be spaced so that the statistical uncertainties between neighboring states be equal;~\cite{pham2012optimal, shenfeld2009minimizing}
Some schemes to adaptively optimize the spacing of intermediate states based on initial exploratory simulations have been proposed~\cite{hayes2017adaptive}. For molecules changing in dense solvent, then the best path is roughly independent of molecule size and shape, so what works for one molecular transformation is likely to be relatively efficient for another~\cite{monroe2004converging}

\subsubsection{Which sampling scheme will work best for my problem?}
\label{sec:sampling_schemes}
Though all alchemical simulations must sample from multiple $\vec{\lambda}$ states, different approaches can be used to achieve this. Fig.~\ref{fig:fig_sampling_scheme} illustrates the four most common schemes. The simplest approach involves running an independent simulation at each of the predefined $\vec{\lambda}$ values (see Fig.~\ref{fig:fig_sampling_scheme} \textbf{A}). This type of scheme is currently used for AMBER TI calculations~\cite{song2019using} and for Siremol/Openmm~\cite{hedges2019biosimspace}. However, if these simulations can be run simultaneously with communication between them, a simple extension allows mixing between these replicas. In this approach, the simulation at each $\vec{\lambda}$ can undergo periodic exchanges with neighboring $\vec{\lambda}$ values. This form of replica exchange (Hamiltonian replica exchange) is based on ideas developed from Monte Carlo simulations of spin glasses by Swendsen and Wang~\cite{swendsen1986replica}. With the Metropolis-Hastings acceptance criterion for exchanges, the generated ensemble of all replicas still samples from the Boltzmann distribution, thus this approach has been used in many different contexts for molecular simulations~\cite{sugita2000multidimensionala,sugita1999replicaexchangea, woods2003developmenta, jiang2010free}. The basic idea of the replica exchange scheme is shown in Fig.~\ref{fig:fig_sampling_scheme} \textbf{B}. It is supported in various software packages that provide alchemical implementations, such as GROMACS~\cite{aldeghi2015accurate}, FEP+~\cite{wang2015accurate}, and NAMD~\cite{jiang2019computing}. A third approach borrows ideas from simulated tempering~\cite{marinari1992simulateda}. In this scheme a single replica rapidly explores all of $\vec{\lambda}$ space by working out optimal weights that allow switching between different intermediate $\vec{\lambda}$ values, as seen in Fig.~\ref{fig:fig_sampling_scheme} \textbf{C} . This approach is also referred to as self adjusted mixture sampling~\cite{lyubartsev1992newa, li2007simulated, tan2017optimally} and while promising, has so far only been supported in OpenMM Tools~\cite{andrearizzi2019choderalab}, and to date has not seen published applications for binding free energy calculations. The last approach makes use of non-equilibrium simulations~\cite{aldeghi2018accurate}. In this approach, only end state $\vec{\lambda}$ replicas ($\vec{\lambda}$=0, $\vec{\lambda}=1$) are simulated at equilibrium; intermediate information is generated from non-equilibrium simulations that rapidly transition between endstates. This approach is available in GROMACS and appears to be coming online in several other packages. A schematic of this approach is shown in Fig.~\ref{fig:fig_sampling_scheme} \textbf{D}.

\begin{figure}
    \includegraphics[width=0.95\columnwidth]{figures/fig8_sampl_scheme/Figure.pdf}
    \caption{\textbf{Four most common sampling strategies are available in different popular software packages.} (\textbf{A}): Multiple replicas in parallel at different lambda states. (\textbf{B}): Hamiltonian replica exchange scheme. (\textbf{C}): Single replica scheme sampling from all $\vec{\lambda}$ states. (\textbf{D}): Non-equilibrium sampling scheme.}
    \label{fig:fig_sampling_scheme}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{figures/fig10_forward_reverse/Figure.pdf}
    \caption{\textbf{Free energy (in k$_{B}$T) for two different relative binding free energy (RBFE) perturbations.} 
    Each plot shows the estimated free energy change using a varying fraction of total simulation time (up to 5 ns total). 
    Subplots (\textbf{A}), (\textbf{C}), and (\textbf{E}) show a three step protocol  for a perturbation involving 3 perturbed atoms, while (\textbf{B}), (\textbf{D}), and (\textbf{F}) shows the same protocol for a perturbation involving 10 perturbed atoms. The first step of the protocol is the recharging then removing van der Waals interactions and then recharging. The difference in energy between the forward (blue) and reverse (red) free energy calculations at the midpoint of the simulation time gives an indication of the overall convergence of the simulation, with differences over 1 k$_{B}$T indicating poor convergence.}
    \label{fig:convergence_forward_reverse}
\end{figure}

Each of these metrics have demonstrated promising results for diagnosing when a simulation has a convergence issue beyond simple convergence of uncertainty estimate. 
Results obtained from calculations with convergence issues should be checked for errors or run for longer before any confidence should be placed in conclusions drawn from their analysis.
In relative calculations that share similar binding modes, for example, and do not induce large conformational changes when in complex with protein, the need to sample exhaustively to converge estimates in free energy differences is often not necessary due to the locality of sampling changes in the molecular topology and shared phase space of the core atoms.
However, even subtly induced changes in protein binding configuration will require more sampling or cause local convergence to a free energy estimate that has high error.
Many enhanced sampling techniques have been proposed to try and overcome barriers in free energy calculations and mitigate convergence to erroneous estimates~\cite{}MISSING citation. 
The confidence a user should have in a free energy estimate is significantly improved when both the uncertainty of the free energy estimate is low, and when other observables have reached a convergence.

Another aspect that can be monitored is bias in the simulation. A simple, but effective, way of checking for bias is to compute the same property of interest by multiple estimators (e.g. free energy difference). See for example the potential mean force with respect to $\vec{\lambda}$ estimated from a bound simulation of a Tyk2 ligand pair of Wang et al.~\cite{wang2015accurate}, as seen in Fig.~\ref{fig:pmf}.
\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{paper/figures/fig_pmf/pmf.png}
    \caption{{\bf Potential of mean force with respect to $\vec{\lambda}$ for TI and MBAR}
    The estimated PMF for a bound calculation of a Tyk2 ligand pair of the Wang et al.~\cite{wang2015accurate} with respect to $\vec{\lambda}$ estimated from TI and MBAR and showing agreement within errorbars. 
    }
    \label{fig:pmf}
\end{figure}

The uncertainty in the free energy, for example, has multiple ways to be estimated, e.g. through standard error propagation methods (including MBAR's estimator, which is based on the same principles as standard error propagation), through bootstrap methods, through multiple independent runs, etc. 
Independent of how the property is estimated, its important to remember that they are \textit{estimations of the property}, not the true underlying property itself. 
These estimators are usually consistent estimators, meaning they will converge to the true answer in the limit of sufficient sampling, not necessarily unbiased ones though.
As such, it is a good idea to subject different estimators to the same data to see if they yield either the same estimate (within error and bias), or if they fluctuate wildly. 
This is not a perfect method as some estimators, such as exponential averaging, will converge significantly more slowly, relative to more accurate estimators like MBAR. 
Therefore, it is a good idea to apply the estimators to different fractions of the data to see if the main estimator of free energy you have chosen is stable.

Each method requires different data from the simulation be collected.  If, for instance, the free energy estimator selected is thermodynamic integration, then values of $\frac{dU}{d\vec{\lambda}}$ at uncorrelated data points must be collected. Once a combination of knowing what type of simulation you will run, which alchemical topology you will simulate, what alchemical path you will simulation along, and what your stopping conditions are, then you are ready to enumerate the information you should capture. Below is a sample of the minimal information you need for a set of common estimators (discussed in more detail in Sec.~\ref{subsec:estimators}):

@@

Free energy information should generally be saved more frequently than coordinate data, approximately at the rate that uncorrelated samples are produced.  
The on-disk size of the data for free energy estimation is often significantly smaller than full atomic coordinates, so the information should be collected frequently. 
However, the information should not be collected \textit{every} time step, as most free energy techniques are operated at equilibrium, and need equilibrated \textit{and decorrelated} samples for an unbiased estimate.
A sample collected every time step will likely result in most samples being discarded due to decorrelation routines in the analysis. However, if it is computationally cheap and disk space is plentiful, do save often. One may safely assume that the correlation time is greater than 100-200 fs even for relatively simple systems such as small molecules in solvent, so saving no more frequently than every 50-100 steps is recommended. 
How decorrelation impacts calculations, and how to compute it is discussed in other section~\ref{sec:decorrelating-samples}.

@@

To deal with this, some practitioners have actually computed relative binding free energies of different binding modes of the same ligand~\cite{palma2012computation}.
For example, a perturbation which adds a methyl to an aromatic ring of a larger ligand might yield one result if the methyl points in one direction, and a different value if it points in the other due to slow ring motions~\cite{lincoff2016comparing}.
One could compute the free energy of turning off the methyl group in one orientation and turning it back on in the other orientation to obtain the free energy difference between the two potential binding modes.
While this approach has precedent, it is relatively difficult to automate at present and requires considerable care.

@@

\paragraph{Computing the timeseries for equilibration detection}
Typically, the timeseries of note $a_t$ analyzed in automated equilibration detection is the negative logarithm of the probability density sampled by the MCMC algorithm (up to an irrelevant additive constant).
For simple independent simulations that sample $x_t \sim \pi(x ; \vec{\lambda})$, this is given by the reduced potential
\begin{eqnarray}
a_t &\equiv& - \ln \pi(x_t; \vec{\lambda}) + c = u(x_t; \vec{\lambda}) .
\end{eqnarray}

@@

\paragraph{Thermodynamic integration (TI)}
By taking the derivative of the free energy with respect to the
variable $\vec{\lambda}$, we find that:
\begin{equation}
\frac{df}{d\vec{\lambda}} = \frac{d}{d\vec{\lambda}} \left[-\ln \int \frac{\exp{-u(\vec{\lambda},\vec{q})}}{Z(\vec{\lambda})} d\vec{q}\right] = \expect{\frac{du(\vec{\lambda},\vec{q})}{d\vec{\lambda}}\
}_{\vec{\lambda}} 
\end{equation}
And than we can then numerically integrate $df/d\vec{\lambda}$ over an alchemical transformation, using a range of different well-established techniques, to obtain:
\begin{equation}
\Delta f    = \int_{0}^{1} \expect{\frac{du(\vec{\lambda},\vec{q})}{d\vec{\lambda}}}_{\vec{\lambda}}  d\vec{\lambda}    
\end{equation}
This approach to calculating the free energy is called thermodynamic integration (TI). Averaging over $\expect{\frac{du}{d\vec{\lambda}}}$ requires fewer uncorrelated samples to reach a given level of relative error
than averaging $e^{-u(\vec{q})}$, as the range is usually narrower with a more Gaussian distribution. Rather than being limited by overlap, as in the case of BAR and MBAR (see below), we are instead limited by the bias in the numerical quadrature, which must be minimized sufficiently to be beneath the level of statistical noise.

Various numerical integration schemes are possible, but the trapezoid
rule provides a simple and robust scheme.  All types of numerical integration can be written as:
\[ \Delta f \approx \sum_{k=1}^{K} w_k
\expect{\frac{du(\vec{\lambda},\vec{q})}{d\vec{\lambda}}}_{k} \] where the weights $w_k$ correspond to a particular choice of numerical integration.
Researchers have tried a large number of different integration schemes~\cite{resat1993studies,jorge2010effect,shyu2009reducing}. However, many other integration choices require specific choices of $\vec{\lambda}$ to minimize bias, which makes them unsuitable when the intermediates
have widely-varying levels of uncertainty. For example, integrating a cubic spline interpolation provided negligible benefits~\cite{paliwal2011benchmark}. For starting researchers, we therefore recommend a simple trapezoid rule scheme, as it allows for maximal flexibility in which values of $\vec{\lambda}$ are simulated.  As fitting to higher order polynomials can have numerical instabilities for some energy functions, and because alternate functional forms might only be appropriate with some types of transformations, expertise and experience is required to perform such numerical integration modifications. In practice, adding 2-3 more intermediate states is typically sufficient to match the performance of these more complicated numerical quadrature schemes.

@@

\paragraph{The multistate Bennett acceptance ratio (MBAR)}
One can generalize Bennett's logic from two states to multiple states to obtain a free energy estimator that uses energy differences between all intermediate states. MBAR gives a system of implicit equations for the free energies $f_i$:
\begin{equation}
f_i = - \ln \sum_{n=1}^{N} \frac{\exp(-u_i(\vec{q}_n))}{\sum_{k=1}^K N_k \exp(f_k-u_k(\vec{q}_n))}
\end{equation}
where there are $N_k$ samples from each of $K$ states, with $\sum_k N_k=N$ the total number of samples. Thus, we need to evaluate the energy function $u_i$ for all samples obtained at all states in the transformation. The equations can be solved by a number of different standard routines. We note that there are only $K-1$ independent equations, so only K-1 of the free energies are independent variables, and one of the $f_i$ must be specified (usually, without loss of generality, setting it to zero).

@@

\section{Checklist}
\label{sec:checklist}
\begin{Checklists*}[p!]
\begin{checklist}{ Know what you want to simulate}
    \textbf{Initial questions you should ask before you set up an alchemical free energy calculation using molecular dynamics simulations}
\begin{itemize}
    \item Do I understand the biology, chemistry and physics of my system?
    \item Have I properly prepared my protein and ligand systems?
    \item Does my system contain any structures that require custom parameters?
    \item What simulation protocol will provide the most evidence to answer my hypothesis?
    \item Are the projected computational expense and runtime realistic for my goals?
    \item Will my protocol be reproducible? 
    \item Will my statistics be reliable? If not, would more replicates solve the problem? 
    \item Can I open-source my data?
\end{itemize}
\end{checklist}

>>>>

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig5_thermo_cyc_abs/Figure_1.pdf}
    \caption{\textbf{Thermodynamic cycle required for an absolute free energy calculation -- absolute free energy of binding example} The fully interacting ligand in water (\textbf{A}), has its charges turned off to pass to (\textbf{B}) followed by turning of van der Waals terms, resulting in a non-interacting ligand in water in (\textbf{C}). Restraints are used on the fully interacting ligand in the binding site of a protein or host molecule (\textbf{D}). The next step is to turn off the charges again (\textbf{E}) followed by the van der Waals interactions resulting in a non-interacting complex state (\textbf{F}). Free energyes can be computed as $\Delta G_{bind} = (\Delta G^{\mathrm{elec}}_{\mathrm{solv}}+ \Delta G^{\mathrm{VdW}}_{\mathrm{solv}})-(\Delta G^{\mathrm{elec}}_{\mathrm{bound}}+ \Delta G^{\mathrm{VdW}}_{\mathrm{bound}}$).
    }
    \label{fig:fig_absolute_thermodynamic_cycle}
\end{figure}

@@

The key consideration for choosing alchemical pathways is that the intermediate states that a given pathway produces, should sample configurational ensembles that change as slowly as possible as $\vec{\lambda}$ changes, while still managing to go from the initial state to the final state as $\vec{\lambda}$ goes from 0 to 1.

@@

\begin{equation}
    U(\vec{r_{ij}},\vec{\lambda}) = 4\epsilon_{ij} \vec{\lambda} \left(\frac{1}{(\alpha(1-\vec{\lambda}) + (r_{ij}/\sigma_{ij})^6)^2} -  \frac{1}{\alpha(1-\vec{\lambda}) + (r_{ij}/\sigma_{ij})^6}\right)
    \label{eq:softcore},
\end{equation}

where $r_{ij}$ is the distance between two particles $i$ and $j$, $\epsilon_{ij}$ and $\sigma_{ij}$ are the Lennard-Jones parameters corresponding to the interaction between particles $i$ and $j$, and $\alpha$ is a constant (0.5 is optimal for the functional form shown above). This functional form has exactly the property we are looking for: it recovers the Lennard-Jones potential when $\vec{\lambda}=1$, and the other endpoint ($\vec{\lambda}=0$), it is exactly zero for all $r_{ij}$ everywhere, and as $\vec{\lambda}$ goes to zero, the $\alpha(1-\vec{\lambda})$ term lowers the infinite energy in the core. There are several different variants of the same functional form~\cite{zacharias1994separationshifted, beutler1994avoiding,pham2011identifying}, but the one given in eq.~\ref{eq:softcore} is easy to understand and implement and fairly numerically stable. This functional form is shown in \textbf{C} and \textbf{D} of Fig.~\ref{fig:fig_what_is_lambda}.

@@

Relative calculations introduce additional choices, such as the order in which to modify nonbonded interactions.
A common process in single topology relative calculations is to first remove electrostatic interactions of any atoms which will be deleted, then modify other non-bonded interactions, then restore electrostatic interactions of any atoms which are being inserted. Although this is a simpler path to understand cognitively and can take advantage of the soft-core potential from eq~\ref{eq:softcore}, this can lead to more intermediate steps and thus be more computationally expensive.
Other schemes, such as simultaneously changing electrostatic and Lennard-Jones interactions with electrostatic ``soft-core'' potentials~\cite{steinbrecher2007nonlinear} may be implemented with fewer intermediate but could require fine-tuning of electrostatic and Lennard-Jones softcore parameters to avoid numerical instabilities. 
At the time of writing, there has not been conclusive evidence to suggest one approach is better in general than the other, so discretion should be left up to the user as to what is viable from both hardware resources, and what the simulation software supports.

A key additional consideration in choosing the alchemical pathway is the choice of spacing of intermediate states.
The spacing depends to some extent on the choice of analysis method, though states should essentially be spaced equidistant in the relevant thermodynamic length~\cite{crooks2007measuringa, sivak2012thermodynamic}.
For BAR/MBAR techniques this means that states should be spaced so that the statistical uncertainties between neighboring states be equal;~\cite{pham2012optimal, shenfeld2009minimizing}
Some schemes to adaptively optimize the spacing of intermediate states based on initial exploratory simulations have been proposed~\cite{hayes2017adaptive}. For molecules changing in dense solvent, then the best path is roughly independent of molecule size and shape, so what works for one molecular transformation is likely to be relatively efficient for another~\cite{monroe2004converging}.

\subsubsection{Which sampling scheme will work best for my problem?}
\label{sec:sampling_schemes}
Though all alchemical simulations must sample from multiple $\vec{\lambda}$ states, different approaches can be used to achieve this. Fig.~\ref{fig:fig_sampling_scheme} illustrates the four most common schemes. The simplest approach involves running an independent simulation at each of the predefined $\vec{\lambda}$ values (see Fig.~\ref{fig:fig_sampling_scheme} \textbf{A}). This type of scheme is currently used for AMBER TI calculations~\cite{song2019using} and for Sire as implemented in BioSimSpace~\cite{hedges2019biosimspace}. However, if these simulations can be run simultaneously with communication between them, a simple extension allows mixing between these replicas. In this approach, the simulation at each $\vec{\lambda}$ can undergo periodic exchanges with neighboring $\vec{\lambda}$ values. This form of replica exchange (Hamiltonian replica exchange) is based on ideas developed from Monte Carlo simulations of spin glasses by Swendsen and Wang~\cite{swendsen1986replica}. With the Metropolis-Hastings acceptance criterion for exchanges, the generated ensemble of all replicas still samples from the Boltzmann distribution, thus this approach has been used in many different contexts for molecular simulations~\cite{sugita2000multidimensionala,sugita1999replicaexchangea, woods2003developmenta, jiang2010free}. The basic idea of the replica exchange scheme is shown in Fig.~\ref{fig:fig_sampling_scheme} \textbf{B}. It is supported in various software packages that provide alchemical implementations, such as GROMACS~\cite{aldeghi2015accurate}, FEP+~\cite{wang2015accurate}, and NAMD~\cite{jiang2019computing}. A third approach borrows ideas from simulated tempering~\cite{marinari1992simulateda}. In this scheme a single replica rapidly explores all of $\vec{\lambda}$ space by working out optimal weights that allow switching between different intermediate $\vec{\lambda}$ values, as seen in Fig.~\ref{fig:fig_sampling_scheme} \textbf{C} . This approach is also referred to as self adjusted mixture sampling~\cite{lyubartsev1992newa, li2007simulated, tan2017optimally} and while promising, has so far only been supported in OpenMM Tools~\cite{andrearizzi2019choderalab}. The last approach makes use of non-equilibrium simulations~\cite{aldeghi2018accurate}. In this approach, only end state $\vec{\lambda}$ replicas ($\vec{\lambda}$=0, $\vec{\lambda}=1$) are simulated at equilibrium; intermediate information is generated from non-equilibrium simulations that rapidly transition between end-states. This approach is available in GROMACS and appears to be coming online in several other packages. A schematic of this approach is shown in Fig.~\ref{fig:fig_sampling_scheme} \textbf{D}.

\begin{figure}
    \includegraphics[width=0.88\columnwidth]{figures/fig8_sampl_scheme/Figure.pdf}
    \caption{\textbf{Four most common sampling strategies.} (\textbf{A}): Multiple replicas in parallel at different lambda states. Each arrow symbolises an independent $\vec{\lambda}$ simulation. (\textbf{B}): Hamiltonian replica exchange scheme. Each arrow represents a short simulation interval before an exchange through Metropolis Hastings acceptance (dice) is attempted. A tick means an accepted exchange a cross a rejected exchange. (\textbf{C}): Single replica scheme sampling from all $\vec{\lambda}$ states. After a short simulation time symbolised by the arrow, the lambda-state is attempted to change until all N lambda states will be sampled. (\textbf{D}): Non-equilibrium sampling scheme, where two equilibrium simulations at the end-states are run as indicated by the blue and pink arrow. Non-equilibrium simulations are attempted at intervals to switch between the two end-states.}
    \label{fig:fig_sampling_scheme}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{figures/fig10_forward_reverse/Figure.pdf}
    \caption{\textbf{Free energy (in k$_{B}$T) for two different relative binding free energy perturbations.} 
    Each plot shows the estimated free energy change using a varying fraction of total simulation time (up to 5 ns total). 
    Subplots (\textbf{A}), (\textbf{C}), and (\textbf{E}) show a three step protocol for a perturbation involving 3 perturbed atoms, while (\textbf{B}), (\textbf{D}), and (\textbf{F}) shows the same protocol for a perturbation involving 10 perturbed atoms. The first step of the protocol is the decharging then removing van der Waals interactions and then recharging. The difference in energy between the forward (blue) and reverse (red) free energy calculations at the midpoint of the simulation time gives an indication of the overall convergence of the simulation, with differences over 1 k$_{B}$T indicating poor convergence.}
    \label{fig:convergence_forward_reverse}
\end{figure}

Each of these metrics have demonstrated promising results for diagnosing when a simulation has a convergence issue beyond simple convergence of uncertainty estimate. 
Results obtained from calculations with convergence issues should be checked for errors or run for longer before any confidence should be placed in conclusions drawn from their analysis.
In relative calculations that share similar binding modes, for example, and do not induce large conformational changes when in complex with protein, the need to sample exhaustively to converge estimates in free energy differences is often not necessary due to the locality of sampling changes in the molecular topology and shared phase space of the core atoms.
However, even subtly induced changes in protein binding configuration will require more sampling or cause local convergence to a free energy estimate that has high error.
Many enhanced sampling techniques have been proposed to try and overcome barriers in free energy calculations and mitigate convergence to erroneous estimates~\cite{xx}. 
The confidence a user should have in a free energy estimate is significantly improved when both the uncertainty of the free energy estimate is low, and when other observables have reached a convergence.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{paper/figures/fig_pmf/pmf.png}
    \caption{{\bf Potential of mean force with respect to $\vec{\lambda}$ for TI and MBAR}
    The estimated PMF for a bound calculation of a Tyk2 ligand pair of the Wang et al.~\cite{wang2015accurate} with respect to $\vec{\lambda}$ estimated from TI and MBAR and showing agreement within errorbars. 
    }
    \label{fig:pmf}
\end{figure}

The uncertainty in the free energy, for example, has multiple ways to be estimated, e.g. through standard error propagation methods (including MBAR's estimator, which is based on the same principles as standard error propagation), through bootstrap methods, through multiple independent runs, etc. 
Independent of how the property is estimated, its important to remember that they are \textit{estimations of the property}, not the true underlying property itself. 
These estimators are usually consistent estimators, meaning they will converge to the true answer in the limit of sufficient sampling, not necessarily unbiased ones though.
As such, it is a good idea to subject different estimators to the same data to see if they yield either the same estimate (within error and bias), or if they fluctuate wildly. See for example the potential of mean force with respect to $\vec{\lambda}$ estimated from a bound simulation of a Tyk2 ligand pair of Wang et al.~\cite{wang2015accurate} for both the MBAR and TI estimators, as seen in Fig.~\ref{fig:pmf}.
This is not a perfect method as some estimators, such as exponential averaging, will converge significantly more slowly, relative to more accurate estimators like MBAR. 
Therefore, it is a good idea to apply the estimators to different fractions of the data to see if the main estimator of free energy you have chosen is stable.

Each method requires different data from the simulation be collected. If, for instance, the free energy estimator selected is thermodynamic integration, then values of $\frac{dU}{d\vec{\lambda}}$ at uncorrelated data points must be collected. Once a combination of knowing what type of simulation you will run, which alchemical topology you will simulate, what alchemical path you will simulate along, and what your stopping conditions are, then you are ready to enumerate the information you should capture. Below is a sample of the minimal information you need for a set of common estimators (discussed in more detail in Sec.~\ref{subsec:estimators}):

@@

Free energy information should generally be saved more frequently than coordinate data, approximately at the rate that uncorrelated samples are produced.  
The on-disk size of the data for free energy estimation is often significantly smaller than full atomic coordinates, so the information should be collected frequently. 
However, the information should not be collected \textit{every} time step, as most free energy techniques are operated at equilibrium, and need equilibrated \textit{and decorrelated} samples for an unbiased estimate.
A sample collected every time step will likely result in most samples being discarded due to decorrelation routines in the analysis. However, if it is computationally cheap and disk space is plentiful, do save often. One may safely assume that the correlation time is greater than 100-200 fs even for relatively simple systems such as small molecules in solvent, so saving no more frequently than every 50-100 steps is recommended. 
How decorrelation impacts calculations, and how to compute it is discussed in Sec.~\ref{sec:decorrelating-samples}.

@@

To deal with this, some practitioners have actually computed relative binding free energies of different binding modes of the same ligand~\cite{palma2012computation}.
For example, a perturbation which adds a methyl to an aromatic ring of a larger ligand might yield one result if the methyl points in one direction, and a different value if it points in the other due to slow ring motions~\cite{lincoff2016comparing, sasmal2020sampling}.
One could compute the free energy of turning off the methyl group in one orientation and turning it back on in the other orientation to obtain the free energy difference between the two potential binding modes.
While this approach has precedent, it is relatively difficult to automate at present and requires considerable care.

@@

\paragraph{Computing the timeseries for equilibration detection}
Typically, the timeseries of note $a_t$ analyzed in automated equilibration detection is the negative logarithm of the probability density ($\pi(x_t; \vec{\lambda}$)) sampled by the MCMC algorithm (up to an irrelevant additive constant).
For simple independent simulations that sample $x_t \sim \pi(x ; \vec{\lambda})$, this is given by the reduced potential
\begin{eqnarray}
a_t &\equiv& - \ln \pi(x_t; \vec{\lambda}) + c = u(x_t; \vec{\lambda}) .
\end{eqnarray}

@@

\paragraph{Thermodynamic integration (TI)}
By taking the derivative of the free energy with respect to the
variable $\vec{\lambda}$, we find that:
\begin{equation}
\frac{df}{d\vec{\lambda}} = \frac{d}{d\vec{\lambda}} \left[-\ln \int \frac{\exp{-u(\vec{\lambda},\vec{q})}}{Z(\vec{\lambda})} d\vec{q}\right] = \expect{\frac{du(\vec{\lambda},\vec{q})}{d\vec{\lambda}}\
}_{\vec{\lambda}} .
\end{equation}
And than we can then numerically integrate $df/d\vec{\lambda}$ over an alchemical transformation, using a range of different well-established techniques, to obtain:
\begin{equation}
\Delta f    = \int_{0}^{1} \expect{\frac{du(\vec{\lambda},\vec{q})}{d\vec{\lambda}}}_{\vec{\lambda}}  d\vec{\lambda}.    
\end{equation}
This approach to calculating the free energy is called thermodynamic integration (TI). Averaging over $\expect{\frac{du}{d\vec{\lambda}}}$ requires fewer uncorrelated samples to reach a given level of relative error
than averaging $e^{-u(\vec{q})}$, as the range is usually narrower with a more Gaussian distribution. Rather than being limited by overlap, as in the case of BAR and MBAR (see below), we are instead limited by the bias in the numerical quadrature, which must be minimized sufficiently to be beneath the level of statistical noise.

Various numerical integration schemes are possible, but the trapezoid
rule provides a simple and robust scheme. All types of numerical integration can be written as:
\[ \Delta f \approx \sum_{k=1}^{K} w_k
\expect{\frac{du(\vec{\lambda},\vec{q})}{d\vec{\lambda}}}_{k} \], where the weights $w_k$ correspond to a particular choice of numerical integration.
Researchers have tried a large number of different integration schemes~\cite{resat1993studies,jorge2010effect,shyu2009reducing}. However, many other integration choices require specific choices of $\vec{\lambda}$ to minimize bias, which makes them unsuitable when the intermediates
have widely-varying levels of uncertainty. For example, integrating a cubic spline interpolation provided negligible benefits~\cite{paliwal2011benchmark}. For starting researchers, we therefore recommend a simple trapezoid rule scheme, as it allows for maximal flexibility in which values of $\vec{\lambda}$ are simulated. As fitting to higher order polynomials can have numerical instabilities for some energy functions, and because alternate functional forms might only be appropriate with some types of transformations, expertise and experience is required to perform such numerical integration modifications. In practice, adding 2-3 more intermediate states is typically sufficient to match the performance of these more complicated numerical quadrature schemes.

@@

\paragraph{The multistate Bennett acceptance ratio (MBAR)}
One can generalize Bennett's logic from two states to multiple states to obtain a free energy estimator that uses energy differences between all intermediate states. MBAR gives a system of implicit equations for the free energies $f_i$:
\begin{equation}
f_i = - \ln \sum_{n=1}^{N} \frac{\exp(-u_i(\vec{q}_n))}{\sum_{k=1}^K N_k \exp(f_k-u_k(\vec{q}_n))},
\end{equation}
where there are $N_k$ samples from each of $K$ states, with $\sum_k N_k=N$ the total number of samples. Thus, we need to evaluate the energy function $u_i$ for all samples obtained at all states in the transformation. The equations can be solved by a number of different standard routines. We note that there are only $K-1$ independent equations, so only K-1 of the free energies are independent variables, and one of the $f_i$ must be specified (usually, without loss of generality, setting it to zero).

@@

\section{Checklist}
\label{sec:checklist}
\begin{Checklists*}
\begin{checklist}{ Know what you want to simulate}
    \textbf{Initial questions you should ask before you set up an alchemical free energy calculation using molecular dynamics simulations}
\begin{itemize}
    \item Do I understand the biology, chemistry and physics of my system?
    \item Have I properly prepared my protein and ligand systems?
    \item Does my system contain any structures that require custom parameters?
    \item What simulation protocol will provide the most evidence to answer my hypothesis?
    \item Are the projected computational expense and runtime realistic for my goals?
    \item Will my protocol be reproducible? 
    \item Will my statistics be reliable? If not, would more replicates solve the problem? 
    \item Can I open-source my data?
\end{itemize}
\end{checklist}
